Field-informed Reinforcement Learning for Learning Large-Scale Collective Tasks

### 2023-05-17 paper.tex
draft discussion

\subsection{Discussion and Results}
\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/start.png}
      \caption{Start}
      \label{fig:initial}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/after.png}
      \caption{After 50 steps}
      \label{fig:after}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/end.png}
      \caption{After 200 steps}
      \label{fig:end}
  \end{subfigure}
  \caption{Qualitative test results of the proposed approach. 
  The agents go towards the phenomenon and try to cover it without collapsing into a single point.
  the grey area represents the area covered by the robots (last figure on the right).}
  \label{fig:simulations}
\end{figure}
\begin{figure}
  \centering
  \hfill
  \begin{subfigure}[b]{0.96\linewidth}
    \includegraphics[width=\linewidth]{imgs/coverage-in-time}
    \caption{Average coverage for each episode in training }
    \label{fig:coverage}
  \end{subfigure}\\
  \hfill
  \begin{subfigure}[b]{0.99\linewidth}
    \includegraphics[width=\linewidth]{imgs/reward-in-time}
    \caption{Reward during training}
    \label{fig:reward}
  \end{subfigure}
  \caption{Training results of field-informed \ac{RL}. 
  Our solution is able to cover the phenomenon better than the baselines and it reaches a higher reward. 
  }
  \label{fig:coverage-reward}
\end{figure}
%% Create a subfigure of coverage-test, inside-test, inside-two-test and coverage-two-test

\begin{figure*}
  \centering
  \begin{subfigure}[b]{0.9\linewidth}
    \includegraphics[width=\linewidth]{imgs/coverage-test.pdf}
    \caption{Coverage in the three types of scenarios}
    \label{fig:coverage-test}
  \end{subfigure}
  \begin{subfigure}[b]{0.9\linewidth}
    \includegraphics[width=\linewidth]{imgs/inside-test}
    \caption{Number of agents inside the phenomenon in the three types of scenarios}
    \label{fig:inside-test}
  \end{subfigure}
  \begin{subfigure}[b]{0.85\linewidth}
    \includegraphics[width=\linewidth]{imgs/inside-two-test}
    \caption{Two zones scenario: agent inside each phenomenon}
    \label{fig:inside-two-test}
  \end{subfigure}
  \begin{subfigure}[b]{0.85\linewidth}
    \includegraphics[width=\linewidth]{imgs/coverage-two-test}
    \caption{Two zones scenario: coverage of each area}
    \label{fig:coverage-two-test}
  \end{subfigure} 
  \caption{Quantative test results in the three types of scenarios.
  We can see that the proposed approach is able to cover and track the phenomenon better then the baselines.
  }
  \label{fig:test}
\end{figure*}
\section{Conclusion and Future Work}
\ga{Page budget: 0.5}
\label{sec:conclusion}

>>>>

\subsection{Discussion and Results}
The results of the training process are summarized in \Cref{fig:training}. 
%
Specifically, we observe that the proposed version achieves higher coverage and total reward compared to other approaches. 
%
Interestingly, despite the global information available in \acp{GNN} without fields, they fail to converge to a good result like the one obtained with the field. 
%
This outcome was expected, as the computed field helps agents encode the necessary information to navigate towards the phenomenon.
%
Furthermore, we note that \ac{GNN} combined with \ac{DQN} and graph replay buffer outperforms the simple \ac{MLP} informed field computation. 
%
This is because relying solely on \ac{MLP} and basic deep learning leads to non-stationary and unstable learning, 
 as evident from the wider confidence interval of the reward over training time.

Focusing now on the test phase, 
 we observe that the field-informed version achieves higher coverage than the other approaches in all scenarios since it shows the ability of our solution to generalize to situations--see \Cref{fig:simulations}. 
% 
We observe that the field-informed version successfully moves the nodes closer to the target phenomenon, 
 distributing them evenly without collapsing into a single central point. 
 \Cref{fig:test} quantitatively presents the results across various previously described scenarios.

The \ac{GNN} versions demonstrate the capability to transfer the learned experience to the test phase whereas the MLP version fails to generalize. 
%
It is worth noting that, in the static configuration, 
 once the desired configuration is achieved in the static case, 
 the nodes cease to move, maintaining the found configuration.
%
Interesting observations arise when we use different scenarios from the training phase. 
 Starting from a division into two zones, 
 we notice that even in this case, the informed version finds a better configuration than the GNN counterpart. 
%
It exhibits both higher overall coverage and manages to divide the system into two equally covered parts. 
%
In contrast, the uninformed version also maintains a fair division but with significantly different coverage between the two parts, 
 indicating an imbalance among the nodes in one of the zones.
%
Finally, the moving scenario emphasizes how the informed version generates a more robust policy for new scenarios. 
 Indeed, we observe that the informed version maintains higher coverage and a greater number of agents on the target phenomenon compared to the other two solutions.
%
The uninformed GNN version, however, 
 fails to generalize its movement behaviour, as evidenced by the simulations where the agents, once reaching the target zone, stop moving due to tracking issues.

In conclusion, the results demonstrate how the proposed idea can generate more robust controllers. 
 By guiding information flow in \acp{GNN}, 
 we improve learning efficiency and alleviate the challenge of encoding relevant information. 
%
Nevertheless, we acknowledge the crucial role of \acp{GNN}. 
 Our modified version of DQN, combined with \acp{GNN}, 
 enables the discovery of robust behaviours in a few episodes, 
 which is challenging to capture with \acp{MLP} combined with \ac{DQN}, even if we use field information.
\begin{figure}
  \centering
  \hfill
  \begin{subfigure}[b]{0.96\linewidth}
    \includegraphics[width=\linewidth]{imgs/coverage-in-time}
    \caption{Average coverage for each episode in training }
    \label{fig:coverage}
  \end{subfigure}\\
  \hfill
  \begin{subfigure}[b]{0.99\linewidth}
    \includegraphics[width=\linewidth]{imgs/reward-in-time}
    \caption{Reward during training}
    \label{fig:reward}
  \end{subfigure}
  \caption{Training results of field-informed \ac{RL}. 
  Our solution is able to cover the phenomenon better than the baselines and it reaches a higher reward. 
  }
  \label{fig:training}
\end{figure}
\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/start.png}
      \caption{Start}
      \label{fig:initial}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/after.png}
      \caption{After 50 steps}
      \label{fig:after}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/end.png}
      \caption{After 200 steps}
      \label{fig:end}
  \end{subfigure}
  \caption{Qualitative test results of the proposed approach. 
  The agents go towards the phenomenon and try to cover it without collapsing into a single point.
  the grey area represents the area covered by the robots (last figure on the right).}
  \label{fig:simulations}
\end{figure}

%% Create a subfigure of coverage-test, inside-test, inside-two-test and coverage-two-test

\begin{figure*}
  \centering
  \begin{subfigure}[b]{0.9\linewidth}
    \includegraphics[width=\linewidth]{imgs/coverage-test.pdf}
    \caption{Coverage in the three types of scenarios}
    \label{fig:coverage-test}
  \end{subfigure}
  \begin{subfigure}[b]{0.9\linewidth}
    \includegraphics[width=\linewidth]{imgs/inside-test}
    \caption{Number of agents inside the phenomenon in the three types of scenarios}
    \label{fig:inside-test}
  \end{subfigure}
  \begin{subfigure}[b]{0.85\linewidth}
    \includegraphics[width=\linewidth]{imgs/inside-two-test}
    \caption{Two zones scenario: agent inside each phenomenon}
    \label{fig:inside-two-test}
  \end{subfigure}
  \begin{subfigure}[b]{0.85\linewidth}
    \includegraphics[width=\linewidth]{imgs/coverage-two-test}
    \caption{Two zones scenario: coverage of each area}
    \label{fig:coverage-two-test}
  \end{subfigure} 
  \caption{Quantative test results in the three types of scenarios.
  We can see that the proposed approach is able to cover and track the phenomenon better than the baselines.
  }
  \label{fig:test}
\end{figure*}
\section{Conclusion and Future Work}
\ga{Page budget: 0.5}
\label{sec:conclusion}
